\section{Conclusion}

This paper presents a dynamic collaborative scheduling system for new energy logistics fleets based on graph algorithms and reinforcement learning. The system models the urban road network as a weighted graph and builds a discrete-time simulation framework that jointly considers task deadlines, vehicle capacity, battery constraints, and charging-station competition. Based on this framework, we implement and compare classical path-planning algorithms, fixed heuristic scheduling rules, and a Q-learning-based hyper-heuristic that dynamically selects dispatching rules according to the system state. Experimental results across different scenario scales show that fixed heuristics are effective in small-scale settings but become increasingly limited as task density and resource competition grow, while the Q-learning hyper-heuristic can learn useful rule-selection patterns and provide competitive performance in dynamic environments. The results also indicate that charging strategy design should be considered together with task scheduling, because battery recovery and charging queues directly affect vehicle availability and final scheduling quality. Overall, this work integrates graph algorithms, simulation modeling, heuristic scheduling, reinforcement learning, and exact optimization into a unified experimental platform, providing a practical basis for further research on new energy logistics fleet scheduling and multi-agent coordination.

\section{Project Reflection}

Through this course project, we gained a deeper understanding of how data structures and algorithms can be applied to a realistic and complex logistics scheduling problem. The project required us to connect graph representation, shortest-path search, task scheduling, simulation design, reinforcement learning, experimental evaluation, and paper writing into a complete system, which was much more challenging than implementing isolated algorithms. During development, we learned that algorithmic performance is not determined by a single module alone; path planning, dispatching rules, battery constraints, charging decisions, and evaluation metrics all interact with each other. We also realized the importance of careful experiment design, reproducible data analysis, and consistent explanation between figures, tables, and text. Although the final system still has room for improvement, such as more stable learning methods and richer state representations, this project helped us improve our ability to design, implement, evaluate, and present a complete algorithmic system, and it gave us a concrete experience of applying data-structure knowledge to real-world intelligent transportation and logistics scenarios.
